---
title: "Exercises Lesson 8"
output: html_notebook
---

This is a practice with Notebooks in R, in which I try to resolve the exercises in Lesson 7 from the Chapman manual. 

I start by loading the needed libraries: 

```{r}
library(car) # to explore pair scatterplot matrices, data explorations,  box-cox transformations
library(corrplot)# for correlation graphs
library(gplots)# for including colors in correlation graphs
library(lattice)# for segment visualizations
library(binom)# for confidence intervals in binomial tests
library(multcomp) # for plotting the means and ci of multigroup comparisons
library(coefplot)# to graph confidence intervals of regression coefficients
library(MCMCpack)# regression with Bayes
library(ggformula) # also for plots
library(dplyr)
library(supernova)



```


I load the dataset

```{r}
hotel.df<- read.csv("https://goo.gl/oaWKgt")
```


1. We explore the data 

```{r}
summary(hotel.df)
describe(hotel.df)
str(hotel.df)
# for cleanliness


```


(1) We also explore the shape of the distributions and the interrelations between variables

```{r}
#for cleanliness 
scatterplotMatrix(hotel.df[,c(1:3,18)])
 # for staff and perks
scatterplotMatrix(hotel.df[,c(4:8,18)])
# for price
scatterplotMatrix(hotel.df[,c(9:12, 18)])
#for environment
scatterplotMatrix(hotel.df[,c(13:15, 18)])
#for the digital benefits
scatterplotMatrix(hotel.df[,c(16:17, 18)])
#for the stay factors
scatterplotMatrix(hotel.df[,c(19,20,22:24, 18)])
```

Most variables are normally distributed. However, we can see that distance traveled, number of nights stayed and  average food spending are rightly skewed. We apply log transformations: :



```{r}
hotel.df$log.dist <- log(hotel.df$distanceTraveled)
hotel.df$log.nights <- log(hotel.df$nightsStayed)
hotel.df$log.sp.food <- log(1+ hotel.df$avgFoodSpendPerNight)

hist(hotel.df$log.dist)
hist(hotel.df$distanceTraveled)

hist(hotel.df$log.nights)
hist(hotel.df$nightsStayed)

hist(hotel.df$log.sp.food)
hist(hotel.df$avgFoodSpendPerNight)

describe(hotel.df)
#for the stay factors
scatterplotMatrix(hotel.df[,c(26:28, 18)])
```


The log transformations do not appear to work well for number of nights and food, so we use box-cox tranformations: 

Since the results of these transformations did not improve the transformations from the log transformations, I will put the code below, but out of the coding area:

(

l.dist<- coef(powerTransform(hotel.df$distanceTraveled))
l.dist
l.nights<- coef(powerTransform(hotel.df$nightsStayed))
l.nights
l.food.sp<- coef(powerTransform(hotel.df$avgFoodSpendPerNight+1))
l.food.sp

hotel.df$cox.dist <- bcPower(hotel.df$distanceTraveled, l.dist)
hotel.df$cox.nights <- bcPower(hotel.df$nightsStayed,l.nights)
hotel.df$cox.sp.food <- bcPower(1+ hotel.df$avgFoodSpendPerNight, l.food.sp)

hist(hotel.df$cox.dist)
hist(hotel.df$cox.nights)
hist(hotel.df$cox.sp.food)
scatterplotMatrix(hotel.df[,c(29:31, 18)])

)
```{r message=TRUE, warning=TRUE, paged.print=TRUE}
# In case the box-cox transformations are desired, just copy the code below

```



2. Exploring further the pattern correlations in the data


```{r}
corrplot.mixed(cor (hotel.df[, c(-21,-25)], use="complete.obs"), upper="ellipse")

corrplot.mixed(cor (hotel.df[, c(1:18,22)], use="complete.obs"),  upper="ellipse")

corrplot.mixed(cor (hotel.df[, c(18:20,22:24,26:28)], use="complete.obs"),  upper="ellipse")

cor(hotel.df[, c(1:18,22)])
```

In this exploration we can see that the satisfaction variables tend to be moderately correlated with each other. The strongest correlations are between the satisfaction with the front staff and other types of satisfaction, including the overall satisfaction (r=.50). Surprisingly, there is a negative correlation between satisfaction with parking price and satisfaction with valet. 

In regards to the remaining contextual variables, average price per night is moderately and negatively correlated with the satisfaction with some prices like dinning and parking (r = .28-.30), and have also negative but more weak correlations with other satisfaction scores, including overall satisfaction (r = -.15). 






3. Correlation coefficients between the items referred to satsisfaction with cleanliness. 

```{r}
#Pearson correlations
cor(hotel.df[, c("satCleanRoom", "satCleanBath", "satCleanCommon")])
#Spearman Correlations
cor(hotel.df[, c("satCleanRoom", "satCleanBath", "satCleanCommon")], method="spearman")
#Polychoric correlations
polychoric(hotel.df[, c("satCleanRoom", "satCleanBath", "satCleanCommon")])
with(hotel.df, polychoric(cbind(satCleanRoom, satCleanBath, satCleanCommon))) #another form

```
 




4. Evaluating if satisfaction with elite membership perks predicts overall satisfaction:

```{r}
m1<- lm(satOverall~satPerks, data=hotel.df)
summary(m1)
plot(m1)
supernova(m1)
gf_point(jitter(satOverall)~jitter(satPerks), data=hotel.df)%>% gf_smooth(satOverall~satPerks)

```





5. We continue evaluating if satisfaction with elite membership perks predicts overall satisfaction, but controlling for the effect of satisfaction with the Front Staff and the city location

```{r}
m2<- lm(satOverall~satPerks + satFrontStaff + satCity, data=hotel.df)
summary(m2)
plot(m2)
supernova(m2)
```





6. We want to know which factor is more important to predict satisfaction with the elite recognition program within the Gold and Platinum members, considering the following factors: front staff, croom cleanliness, the points that we award, and the membership perks given. 

```{r}
par(mfrow =c(2,2))
m3<- lm(satRecognition~ satCleanRoom+ satFrontStaff + satPerks +satPoints , data= subset(hotel.df, eliteStatus %in% c("Gold", "Platinum")))
summary(m3)
plot(m3)
supernova(m3)
coefplot(m3, intercept =FALSE, outerCI=1.96, LwdOUTER =1.5,
         ylab="Factors",
         xlab="Association with Satisfaction in the Elite Program")
```


7. It is important to consider that this is not an exact measurement of the contribution of each factor to the satisfaction. To get more valid measures we will actually have to measure real changes in those factors that would be derived from the money increase. The measures we have are based on how prior variations in satisfaction with each factor are related to satisfaction, but we are not sure if those variations are in line with the ones tha we will create with the investments. Yet, it can work as an approximation. 

8. Considering these results, it looks like cleanliness is not a major factor to increase the satisfaction with the membership program. 


9. Now, if I understand correct, we want to understand the relationship of the money spent in food with elite status and satisfaction with food price


```{r}
hotel.df$eliteStatus<-factor(hotel.df$eliteStatus)
par(mfrow =c(2,2))
m4<- lm(log.sp.food~ eliteStatus + satDiningPrice , data= hotel.df)
summary(m4)
plot(m4)
supernova(m4)
coefplot(m4, intercept =FALSE, outerCI=1.96, LwdOUTER =1.5,
         ylab="Factors",
         xlab="Association with Money Spent in Food by Clients")
```
Just for comparison, I want to see the fit of the model if we did not use the logarithmic transformation for price of the food

```{r}

par(mfrow =c(2,2))
m5<- lm(avgFoodSpendPerNight~ eliteStatus + satDiningPrice , data= hotel.df)
summary(m5)
plot(m5)
supernova(m5)
coefplot(m5, intercept =FALSE, outerCI=1.96, LwdOUTER =1.5,
         ylab="Factors",
         xlab="Association with Money Spent in Food by Clients")
```

Now I want to make sure there are not important interactions between the elite status and satisfaction with food to predict higher spending in food

```{r}

par(mfrow =c(2,2))
m6<- lm(log.sp.food~ eliteStatus + satDiningPrice + eliteStatus:satDiningPrice , data= hotel.df)
summary(m6)
plot(m6)
supernova(m6)
coefplot(m6, intercept =FALSE, outerCI=1.96, LwdOUTER =1.5,
         ylab="Factors",
         xlab="Association with Money Spent in Food by Clients")
```



10. I am really confused with these questions, it looks like satDinningPrice represents satisfaction with dinner? or satisfaction with the dinner price???

```{r}
m7<- lm(log.sp.food ~  satDiningPrice, data= hotel.df)
summary(m7)

m8<- lm(satDiningPrice ~ log.sp.food , data= hotel.df)
summary(m8)
```
11. Plot how the food spent per night can be predicted by the number of nights: 

```{r}
par(mfrow =c(2,2))
m9<- lm(log.sp.food~ log.nights , data= hotel.df)
summary(m9)
plot(m9)
supernova(m9)
gf_point(log.sp.food~ log.nights , data= hotel.df) %>%gf_smooth(log.sp.food~ log.nights )

gf_point(exp(log.sp.food)~ exp(log.nights) , data= hotel.df) %>%gf_smooth()

gf_point(avgFoodSpendPerNight~nightsStayed , data= hotel.df) %>%gf_smooth()

gf_point(jitter(exp(log.sp.food))~ jitter(exp(log.nights)) , data= hotel.df) %>%gf_smooth( )

#prediction for someone who expend 40 nights
exp(2.64987+(.20152*log(40)))

#Spearman correlation to see if results go in line with the log model or the real unit model


with(hotel.df, cor(cbind(avgFoodSpendPerNight,nightsStayed), method="spearman"))
with(hotel.df, cor(cbind(log.sp.food,log.nights)))
```






Seeing if the difference is the difference between nights spent and spending on food is different for platinum members

```{r}
hotel.df$elite.p <- ifelse(hotel.df$eliteStatus=="Platinum", 1,0)

par(mfrow =c(2,2))
m10<- lm(log.sp.food~ log.nights + elite.p + elite.p:log.nights , data= hotel.df)
summary(m10)
plot(m10)
supernova(m10)

coefplot(m10, intercept =FALSE, outerCI=1.96, LwdOUTER =1.5,
         ylab="Factors",
         xlab="Association with Money Spent in Food by Clients")
```


The manual used a different approach to answer this question, by looking at the model in the platinum subset, and comparing the relation with the elite status

```{r}
par(mfrow =c(2,2))

plat.df <- subset(hotel.df, eliteStatus =="Platinum")

m11<- lm(log.sp.food~ log.nights , data= plat.df)
summary(m11)
plot(m11)
supernova(m11)

#To see the scatterplot segregated by Platinum members versus non Platinum members
gf_point(jitter(exp(log.sp.food))~ jitter(exp(log.nights)) , col= "red", data= subset(hotel.df, eliteStatus=="Platinum")) %>% gf_point(jitter(exp(log.sp.food))~ jitter(exp(log.nights)) , col= "blue", data= subset(hotel.df, eliteStatus !="Platinum"))



#To compare the two models grafically, the one for the whole population versus the model for the platinum population

hotel.df$predictedm9<- fitted(m9)
plat.df$predictedm11<- fitted(m11)


gf_point(hotel.df$log.sp.food~ fitted(m9) , col= "red", ylab="money spent($)", xlab="Predicted scores") %>%gf_point(log.sp.food~ predictedm11 , data=plat.df, col= "blue") 

```
gf_point(hotel.df$log.sp.food~ fitted(m9) , col= "red", ylab="money spent($)", xlab="Predicted scores in all")


%>% gf_point(jitter(exp(log.sp.food))~ jitter(exp(log.nights)) , col= "blue", data= subset(hotel.df, eliteStatus !="Platinum"))




13. Similar to exercise 6,we want to know which factor is more important to predict satisfaction with the elite recognition program within the Gold and Platinum members, considering the following factors: front staff, croom cleanliness, the points that we award, and the membership perks given. However, now we are going to use Bayesian regression. 

To remember the prior approach: 
```{r}
#To remember the prior approach: 
par(mfrow =c(2,2))
m3<- lm(satRecognition~ satCleanRoom+ satFrontStaff + satPerks +satPoints , data= subset(hotel.df, eliteStatus %in% c("Gold", "Platinum")))
summary(m3)
plot(m3)
supernova(m3)
coefplot(m3, intercept =FALSE, outerCI=1.96, LwdOUTER =1.5,
         ylab="Factors",
         xlab="Association with Satisfaction in the Elite Program")


```


And the Bayesian regression: 

```{r}
#The new approach with Bayes: 
par(mfrow =c(2,2))
m3.b<- MCMCregress(satRecognition~ satCleanRoom+ satFrontStaff + satPerks +satPoints , data= subset(hotel.df, eliteStatus %in% c("Gold", "Platinum")))

summary(m3.b)



```

14. Comparamos los coeficientes de regresión de los dos modelos

```{r}
hotel.rec.compare <- data.frame(classical = coef(m3)[-1], 
                                bayesian  = summary(m3.b)$statistics[c(-1, -6), 1])
hotel.rec.compare
plot(bayesian ~ classical, data=hotel.rec.compare)
abline(0,1)
cor(hotel.rec.compare$classical, hotel.rec.compare$bayesian)
```





